---
title: "Prediction"
author: "Cania Chung"
date: "11/6/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
```

## Practical Machine Learning Prediction Writeup

### Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Loading Data

```{r loaddata}
#
# Data
# The training data for this project 
#
TrainURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
TrainFile <- "data/pml-training.csv"

if (!file.exists("data")) {
    dir.create("data")
}

if (!file.exists(TrainFile)) {
    download.file(TrainURL, destfile=TrainFile)
}

training <- read.csv(TrainFile)

#
# The test data for this project
#
TestURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
TestFile <- "data/pml-testing.csv"

if (!file.exists(TestFile)) {
    download.file(TestURL, destfile=TestFile)
}

testing <- read.csv(TestFile)

#
# Create a partition using package caret. Teh Training Dataset is 70:30 ratio.
# 
inTrain <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain,]
TestSet <- training[-inTrain,]
dim(TrainSet)
dim(TestSet)
```

## Cleaning Data

Datasets have 160 variables. In this section, remove NA, Near-zero variables and the identity variables.

``` {r clean}
# First, remove the Near Zero Variables
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet <- TestSet[,-NZV]
dim(TrainSet)
dim(TestSet)

# Second, remove NA variables
ALLNA <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet [,ALLNA==FALSE]
TestSet <- TestSet[,ALLNA==FALSE]
dim(TrainSet)
dim(TestSet)

# Third, remove the identity variables
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainSet)
dim(TestSet)
```

## Prediction
### Random Forests
```{r RandomForest}
set.seed(12345)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFitRandForest <- train(classe ~ ., data=TrainSet, method="rf",trControl=controlRF)
modFitRandForest$finalModel

## Validate on Test Data
predictRandForest <- predict(modFitRandForest, newdata=TestSet)
confMatRandForest <- confusionMatrix(predictRandForest, TestSet$classe)
confMatRandForest

# plot matrix results
plot(confMatRandForest$table, col = confMatRandForest$byClass, 
     main = paste("Random Forest - Accuracy =",
                  round(confMatRandForest$overall['Accuracy'], 4)))


```

### Decision Tree

```{r DecisionTree}
set.seed(12345)
modFitDecTree <- rpart(classe ~ ., data=TrainSet, method="class")
fancyRpartPlot(modFitDecTree)

# prediction on Test dataset
predictDecTree <- predict(modFitDecTree, newdata=TestSet, type="class")
confMatDecTree <- confusionMatrix(predictDecTree, TestSet$classe)
confMatDecTree


# plot matrix results
plot(confMatDecTree$table, col = confMatDecTree$byClass, 
     main = paste("Decision Tree - Accuracy =",
                  round(confMatDecTree$overall['Accuracy'], 4)))

```

### Generalized Boosted Model

```{r gbm}
# model fit
set.seed(12345)
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=TrainSet, method = "gbm", trControl = controlGBM, verbose = FALSE)
modFitGBM$finalModel

# prediction on Test dataset
predictGBM <- predict(modFitGBM, newdata=TestSet)
confMatGBM <- confusionMatrix(predictGBM, TestSet$classe)
confMatGBM

# plot matrix results
plot(confMatGBM$table, col = confMatGBM$byClass, 
     main = paste("GBM - Accuracy =", round(confMatGBM$overall['Accuracy'], 4)))

```

## Use Random Forest prediction to apply to the test data

```{r applyPredictTest}
predictTEST <- predict(modFitRandForest, newdata=testing)
predictTEST
```
